---
title: 
author: "Jessie Tian"
date: 
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(rms)
library(lattice)
library(dplyr)
library(tidyr)
library(survival)
library(ggplot2)
library(skimr)
library(naniar)
library(survival)
library(survminer)
library(lubridate)
library(mice)
library(gtsummary)
library(dplyr)

require("knitr")
```


```{r}

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
stopvape <- read_csv("unimputed_dataset_allages.csv", col_names = TRUE)

stopvape.dd <- datadist(stopvape)
options(datadist="stopvape.dd")

stopvape_complete <- na.omit(stopvape)
sum(is.na(stopvape_complete))


#Complete cases 
nrow(stopvape[!complete.cases(stopvape), ])/nrow(stopvape)*100

#Incomplete cases 17.82%
stopvape_comp <- stopvape[complete.cases(stopvape),]
# df_comp$wave <- NULL
#n=5,288, p=37



```


ecusv_3: Confidence in quitting vaping within next month on a scale of 1-10
ecu17_1: Intention to quit at baseline (0= yes, but within next 6 months or beyond 6 months, 1= within next month,)
ecusv_2: Past-year quit attempts (0= no, 1=yes)
mhealth1_1: Perceived mental heath status (0= fair or poor, 1= good, very good, or excellent)

# prepare data
```{r}
# Convert categorical variables to a factor with proper labels
stopvape$gender_adj <- factor(stopvape$gender,
                            levels = c(0, 1, 2),
                            labels = c("Men", "Women", "Others"))
stopvape$sexorient_adj <- factor(stopvape$sexorient,
                               levels = c(0, 1, 2),
                               labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed"))

stopvape <- stopvape %>%
  mutate(age = ifelse(nchar(as.character(age)) > 3, NA, age))


summary(stopvape$age)
```

# Descriptive table
```{r}

#split into groups low and high quitting confidence 
stopvape$conf_group <- cut(stopvape$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right = TRUE)

# Get descriptive statistics for unique user IDs only
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n}  ({p}%)"
    )
  )

# Filter to unique user IDs, create confidence groups, and get descriptive statistics
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  tbl_summary(
    by = conf_group,  # Group by confidence level
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%        # Add p-values for group comparisons
  add_overall()      # Add overall column


# Select specific variables and group by confidence
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  select(age,gender,sexorient,race,status, Time.elapsed, ecu17_1, ecusv_2, ecusv_3, mhealth1_1, conf_group) %>%
  tbl_summary(
    by = conf_group,
    statistic = list(
      all_continuous() ~ "{median} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%
  add_overall()

# Calculate statistics using unique user_id
quit_stats <- stopvape %>%
  group_by(userid) %>%
  summarise(
    quit_attempts = n(),  # Count records per user (each record = 1 attempt)
    .groups = 'drop'
  )

# Calculate final statistics
final_stats <- quit_stats %>%
  summarise(
    # Average number of quit attempts per participant
    avg_quit_attempts = mean(quit_attempts),
    # Number of participants with only one challenge
    participants_one_challenge = sum(quit_attempts == 1),
    # Maximum number of challenges completed
    max_challenges = max(quit_attempts)
  )

print(final_stats)


# Count how many times each userid appears (number of challenges per user)
user_challenge_counts <- table(stopvape$userid)

```
```{r}
episode_counts <- stopvape %>%
  group_by(userid) %>%
  summarise(
    n_episodes = n(),
    .groups = "drop"
  )

summary(episode_counts$n_episodes)

episode_distribution <- episode_counts %>%
  count(n_episodes) %>%
  mutate(
    percent = n / sum(n) * 100
  )

episode_distribution

repeat_summary <- episode_counts %>%
  summarise(
    total_participants = n(),
    repeat_participants = sum(n_episodes > 1),
    repeat_percent = repeat_participants / total_participants * 100
  )

repeat_summary


repeat_contribution <- stopvape %>%
  left_join(episode_counts, by = "userid") %>%
  summarise(
    total_episodes = n(),
    episodes_repeat = sum(n_episodes > 1),
    percent_repeat = episodes_repeat / total_episodes * 100
  )

repeat_contribution

library(ggplot2)

ggplot(episode_counts, aes(x = n_episodes)) +
  geom_bar() +
  labs(
    x = "Number of Quit Challenges per Participant",
    y = "Number of Participants",
    title = "Distribution of Quit Challenge Episodes per Participant"
  )



```


# Missing data and multiple imputations (pattern of missingness)
```{r}

skim(stopvape)
summary(stopvape, na.rm = TRUE) 
stopvape$userid <- as.factor(stopvape$userid)

selected_vars <- stopvape %>% select(Time.elapsed, status, ecusv_3, age, gender_adj)

# Select the variables of interest
selected_vars <- stopvape %>% 
  select(Time.elapsed, status, ecusv_3, age, gender_adj, ecu17_1, ecusv_2, mhealth1_1)

# Create a summary of missing values
missing_summary <- selected_vars %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%  # count missing per column
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Missing_Count") %>%
  mutate(Total = nrow(stopvape),
         Missing_Percent = round((Missing_Count / Total) * 100, 2))

missing_summary



# Explore patterns of missing data
na.pat <- naclus(stopvape[,-1])
naplot(na.pat)

vis_miss(stopvape) #display entire dataset to find missing, if sort by time, can see if time dependent missingness throughout study
gg_miss_var(stopvape) # give number missing sorted
gg_miss_var(stopvape, show_pct = TRUE) # shown as percentage
gg_miss_case(stopvape)

# gg_miss_upset(stopvape) # see what is missing when other variables are missing
n_var_miss(stopvape) #max number missing is 23 variables

ggplot(
  data = stopvape,
  mapping = aes(x = Time.elapsed, y = ecusv_3)) +     
  geom_miss_point()

stopvape$age[stopvape$age== 5198023886] <- NA

### Generate 5 imputed data sets

(v <- names(stopvape)[c(2,14, 32, 33, 43, 44, 42, 12, 13, 28, 40, 41)])

# md.pattern(stopvape[,v], rotate.names = TRUE)

set.seed(3098) 
stopvape.mice <- mice(stopvape[,v], m = 5, method = "cart") # large number of unbalanced factor variables, use cart method 
stopvape.mice$method["userid"] <- ""

# Prevent userid from being imputed

stopvape.mice$method
plot(stopvape.mice)

densityplot(stopvape.mice) # for continuous variables
densityplot(stopvape.mice, ~ ecusv_3|.imp) #look at self-confidence by imputation


```



```{r}
# Unadjusted model
fit1 <- with(
  stopvape.mice,
  coxph(
    Surv(Time.elapsed, status) ~ ecusv_3 +
    cluster(userid)
  )
)
pooled_fit1 <- pool(fit1)

summary(
  pooled_fit1,
  conf.int = TRUE,
  exponentiate = TRUE
)

# Adjusted model

fit2 <- with(
  stopvape.mice,
  coxph(
    Surv(Time.elapsed, status) ~
      ecusv_3 +
      age +
      country +
      gender_adj +
      sexorient_adj +
      race +
      ecu17_1 +
      ecusv_2 +
      mhealth1_1 +
      cluster(userid)
  )
)

pooled_fit2 <- pool(fit2)

summary(
  pooled_fit2,
  conf.int = TRUE,
  exponentiate = TRUE
)


# Interaction model (gender × quitting confidence)

fit3 <- with(
  stopvape.mice,
  coxph(
    Surv(Time.elapsed, status) ~
      ecusv_3 * gender_adj +
      age +
      country +
      sexorient_adj +
      race +
      ecu17_1 +
      ecusv_2 +
      mhealth1_1 +
      cluster(userid)
  )
)

pooled_fit3 <- pool(fit3)

summary(
  pooled_fit3,
  conf.int = TRUE,
  exponentiate = TRUE
)

# Demographic adjustment model

fitdemo <- with(
  stopvape.mice,
  coxph(
    Surv(Time.elapsed, status) ~
      ecusv_3 +
      age +
      country +
      gender_adj +
      sexorient_adj +
      race +
      cluster(userid)
  )
)

pooled_fitdemo <- pool(fitdemo)

summary(
  pooled_fitdemo,
  conf.int = TRUE,
  exponentiate = TRUE
)

#########################
# Adjusted model

fit_strata <- with(
  stopvape.mice,
  coxph(
    Surv(Time.elapsed, status) ~
      ecusv_3 +
      age +
      strata(country) +
      strata(gender_adj) +
      strata(sexorient_adj) +
      race +
      ecu17_1 +
      strata(ecusv_2) +
      mhealth1_1 +
      cluster(userid)
  )
)

pooled_fit_strata <- pool(fit_strata)

summary(
  pooled_fit_strata,
  conf.int = TRUE,
  exponentiate = TRUE
)
```
Survival probability for main model 
```{r}
###################################################
### Survival probabilities at 24, 48, 72 hours
### by confidence group (Low vs High)
### across all imputed datasets
###################################################

time_points <- c(1, 2, 3)  # 1, 2, 3 days = 24, 48, 72 hours

surv_probs <- lapply(1:stopvape.mice$m, function(i) {
  data_i <- complete(stopvape.mice, i)
  
  # Create confidence group in each imputed dataset
  data_i$conf_group <- cut(data_i$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right  = TRUE)
  
  # Fit KM separately for each confidence group
  km_fit <- survfit(Surv(Time.elapsed, status) ~ conf_group,
                    data = data_i)
  
  # Extract survival probabilities at each time point
  km_summary <- summary(km_fit, times = time_points)
  
  data.frame(
    time       = km_summary$time,
    surv       = km_summary$surv,
    lower      = km_summary$lower,
    upper      = km_summary$upper,
    conf_group = km_summary$strata,
    imp        = i
  )
})

surv_probs_df <- do.call(rbind, surv_probs)

# Clean up strata label
surv_probs_df$conf_group <- gsub("conf_group=", "", surv_probs_df$conf_group)

cat("\n-- Raw survival probabilities per imputation --\n")
print(surv_probs_df)


#################
###################################################
### Pool using Rubin's rules on log(-log(S)) scale
### This is the recommended scale for pooling
### survival probabilities (more stable than
### pooling S directly)
###################################################

surv_pooled <- surv_probs_df %>%
  group_by(time, conf_group) %>%
  summarise(
    # Rubin's rules on log(-log(S)) scale
    m          = n(),
    
    # Transform to log(-log(S)) for pooling
    lls        = mean(log(-log(surv))),
    lls_var    = var(log(-log(surv))),
    
    # Within-imputation variance approximation
    # using delta method on log(-log(S))
    mean_surv  = mean(surv),
    sd_surv    = sd(surv),
    
    # Simple average (acceptable approximation)
    surv_avg   = mean(surv),
    lower_avg  = mean(lower),
    upper_avg  = mean(upper),
    
    .groups = "drop"
  ) %>%
  mutate(
    # Back-transform pooled log(-log(S)) to survival
    surv_pooled = exp(-exp(lls)),
    
    # Convert time back to hours for reporting
    time_hours  = time * 24,
    
    # Format as percentage
    surv_pct    = round(surv_avg * 100, 1),
    lower_pct   = round(lower_avg * 100, 1),
    upper_pct   = round(upper_avg * 100, 1)
  )

cat("\n-- Pooled survival probabilities (Low vs High self-efficacy) --\n")
print(surv_pooled %>% 
        select(time_hours, conf_group, surv_pct, lower_pct, upper_pct))

```

```{r}

schoenfeld_main <- lapply(1:stopvape.mice$m, function(i) {
  data_i <- complete(stopvape.mice, i)
  fit_i <- coxph(Surv(Time.elapsed, status) ~ ecusv_3 + age + country + gender_adj + sexorient_adj + race + ecu17_1 + ecusv_2 + mhealth1_1,
                 data = data_i)
  zph_i <- cox.zph(fit_i)
  data.frame(
    term    = rownames(zph_i$table),
    chisq   = zph_i$table[, "chisq"],
    p_value = zph_i$table[, "p"],
    imp     = i
  )
})

schoenfeld_main_df <- do.call(rbind, schoenfeld_main)

# Summary: mean p-value and proportion significant across imputations
schoenfeld_main_summary <- schoenfeld_main_df %>%
  group_by(term) %>%
  summarise(
    mean_chisq  = mean(chisq),
    mean_pvalue = mean(p_value),
    prop_sig    = mean(p_value < 0.05),
    .groups = "drop"
  )
cat("\n-- Schoenfeld PH Test: Main Dataset --\n")
print(schoenfeld_main_summary)

# Plot: imputation 1 as representative
data_1 <- complete(stopvape.mice, 1)
fit_1 <- coxph(Surv(Time.elapsed, status) ~ ecusv_3 + age + gender_adj +
                 ecu17_1 + ecusv_2 + mhealth1_1,
               data = data_1)
zph_1 <- cox.zph(fit_1)

par(mfrow = c(2, 3))
plot(zph_1, col = "blue", lwd = 2)
par(mfrow = c(1, 1))

# ecusv_3 specifically
plot(zph_1[1],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)
plot(zph_1[2],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)

plot(zph_1[3],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)

plot(zph_1[4],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)

plot(zph_1[5],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)

plot(zph_1[6],
     main = "Schoenfeld Residuals — ecusv_3 (Main, Imp 1)")
abline(h = coef(fit_1)["ecusv_3"], col = "red", lwd = 2, lty = 3)




```




# Univariable Kaplan-Meier curve of vaping relapse survival by quitting confidence levels (low, high)

```{r}

# no missing values in self-efficacy variable
sum(is.na(stopvape$ecusv_3))

# Create confidence group on unimputed data
stopvape_km <- stopvape %>%
  filter(!is.na(ecusv_3)) %>%   # remove only if NAs exist; safe to keep regardless
  mutate(conf_group = cut(ecusv_3,
                          breaks = c(0, 5, 10),
                          labels = c("Low", "High"),
                          right  = TRUE))

# Fit KM
surv_unimputed <- survfit(Surv(Time.elapsed, status) ~ conf_group,
                          data = stopvape_km)

# --- Time point: 48 hours = 2 days ---
summary(surv_unimputed, times = 2)

# --- Time point: 7 days ---
summary(surv_unimputed, times = 7)

# Plot KM curve
km_plot <- ggsurvplot(
  surv_unimputed,
  data       = stopvape_km,
  conf.int   = TRUE,
  pval = TRUE,
  pval.coord = c(38, 0.3),
  risk.table = TRUE,
  xlim       = c(0, 45),
  break.time.by = 7,
  xlab       = "Time Elapsed (days)",
  ylab       = "Probability of Remaining Vape-Free",
  title      = "KM Curve by Self-Efficacy Group",
  legend.labs = c("Low Self-Efficacy", "High Self-Efficacy"),
  palette    = c("steelblue", "red3")
)

# Save as PNG
ggsave(
  filename = "km_curve_unimputed.png",
  plot = km_plot$plot,   # IMPORTANT: extract the ggplot component
  width = 8,
  height = 6,
  dpi = 300
)
```


### log-rank MH test → highly significant difference between the survival curves of the groups
```{r}
#log-rank MH test 
#compute the test of the difference between the survival curves
sur.mh <- survdiff(Surv(Time.elapsed, status) ~ conf_group, data = stopvape)
sur.mh 

```


Report the proportion of episodes that are censored due to inactivity/last-entry rules and summarize engagement patterns over time.

stopvape_firstquit: This is the dataset containing the first quit challenges  for each participant (n=389). 

stopvape_3moodcraving: This is the dataset containing at least 3 mood or craving ratings per challenges. It contains 126 challenge data of 111 participants.
```{r}
stopvape_firstquit <- read_csv("unimputed_dataset_allages_firstquit(in).csv", col_names = TRUE)

# prepare data
# Convert categorical variables to a factor with proper labels
stopvape_firstquit $gender_adj <- factor(stopvape_firstquit$gender,
                            levels = c(0, 1, 2),
                            labels = c("Men", "Women", "Others"))
stopvape_firstquit $sexorient_adj <- factor(stopvape_firstquit$sexorient,
                               levels = c(0, 1, 2),
                               labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed"))
stopvape_firstquit <- stopvape_firstquit %>%
  mutate(age = ifelse(nchar(as.character(age)) > 3, NA, age))



summary(stopvape_firstquit)

#split into groups low and high quitting confidence 
stopvape_firstquit$conf_group <- cut(stopvape_firstquit$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right = TRUE)

```

```{r}
(v <- names(stopvape_firstquit )[c(2, 14, 32, 33, 43, 44, 42, 12, 13, 28, 40, 41)])
set.seed(3098) 
stopvape_firstquit.mice <- mice(stopvape_firstquit[,v], method = "cart") # large number of unbalanced factor variables, use cart method 
stopvape_firstquit.mice$method
plot(stopvape_firstquit.mice)

densityplot(stopvape_firstquit.mice) # for continuous variables
densityplot(stopvape_firstquit.mice, ~ ecusv_3|.imp) #look at self-confidence by imputation



```

```{r}
# return all imputed datasets combined into one big data frame
stopvape_firstquit.pooled <- complete(stopvape_firstquit.mice, method = "long") 
summary (stopvape_firstquit.pooled)
nrow(stopvape_firstquit.pooled)

surv_elapsed <- Surv(time = stopvape_firstquit.pooled$Time.elapsed, 
                 event = stopvape_firstquit.pooled$status)


fit1_firstquit <- coxph(surv_elapsed~ecusv_3,cluster(stopvape_firstquit.pooled$userid),data=stopvape_firstquit.pooled)
fit2_firstquit  <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race+ecu17_1+ecusv_2+mhealth1_1,cluster(stopvape_firstquit.pooled$userid),data=stopvape_firstquit.pooled)
fit3_firstquit  <- coxph(surv_elapsed~ecusv_3*gender_adj,cluster(stopvape_firstquit.pooled$userid),data=stopvape_firstquit.pooled)

fitdemo_firstquit  <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race,cluster(stopvape_firstquit.pooled$userid),data=stopvape_firstquit.pooled)


fit1_firstquit
fit2_firstquit
fit3_firstquit
fitdemo_firstquit

#95% CI
exp(confint(fit1_firstquit))
exp(confint(fit2_firstquit))
exp(confint(fit3_firstquit))
exp(confint(fitdemo_firstquit))

#Run AIC to compare predictive performance of models
AIC(fit1_firstquit)
AIC(fit2_firstquit)
AIC(fit3_firstquit)
AIC(fitdemo_firstquit)


```


# Sensitvity analysis: Challenge episodes with 3 cravings or more 
# prepare data
```{r}

stopvape_3moodcraving <-read_csv("unimputed_dataset_allages_3moodcraving(in).csv", col_names = TRUE)


# Convert categorical variables to a factor with proper labels
stopvape_3moodcraving $gender_adj <- factor(stopvape_3moodcraving$gender,
                            levels = c(0, 1, 2),
                            labels = c("Men", "Women", "Others"))
stopvape_3moodcraving $sexorient_adj <- factor(stopvape_3moodcraving$sexorient,
                               levels = c(0, 1, 2),
                               labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed"))
stopvape_3moodcraving <- stopvape_3moodcraving %>%
  mutate(age = ifelse(nchar(as.character(age)) > 3, NA, age))



summary(stopvape_3moodcraving)

#split into groups low and high quitting confidence 
stopvape_3moodcraving$conf_group <- cut(stopvape_3moodcraving$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right = TRUE)

# Filter to unique user IDs, create confidence groups, and get descriptive statistics
stopvape_3moodcraving %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  tbl_summary(
    by = conf_group,  # Group by confidence level
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%        # Add p-values for group comparisons
  add_overall()      # Add overall column
```

```{r}
(v <- names(stopvape_3moodcraving)[c(2, 14, 32, 33, 43, 44, 42, 12, 13, 28, 40, 41)])
set.seed(3098) 
stopvape_3moodcraving.mice <- mice(stopvape_3moodcraving[,v], method = "cart") 

stopvape_3moodcraving.mice$method
plot(stopvape_3moodcraving.mice)

densityplot(stopvape_3moodcraving.mice) # for continuous variables
densityplot(stopvape_3moodcraving.mice, ~ ecusv_3|.imp) #look at self-confidence by imputation

```

```{r}
fit1_3moodcraving <- with(
  stopvape_3moodcraving.mice,
  coxph(Surv(Time.elapsed, status) ~ ecusv_3 + cluster(userid))
)

fit2_3moodcraving <- with(
  stopvape_3moodcraving.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 + age + country + gender_adj +
          sexorient_adj + race + ecu17_1 +
          ecusv_2 + mhealth1_1 +
          cluster(userid))
)

fit3_3moodcraving <- with(
  stopvape_3moodcraving.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 * gender_adj +
          cluster(userid))
)

fitdemo_3moodcraving <- with(
  stopvape_3moodcraving.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 + age + country +
          gender_adj + sexorient_adj + race +
          cluster(userid))
)

pooled_fit1_3moodcraving <- pool(fit1_3moodcraving)
pooled_fit2_3moodcraving <- pool(fit2_3moodcraving)
pooled_fit3_3moodcraving <- pool(fit3_3moodcraving)
pooled_fitdemo_3moodcraving <- pool(fitdemo_3moodcraving)

summary(pooled_fit1_3moodcraving, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fit2_3moodcraving, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fit3_3moodcraving, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fitdemo_3moodcraving, conf.int = TRUE, exponentiate = TRUE)

```



```{r}
stopvape_relapsed <- stopvape %>%
  filter(status != 0) %>%
  mutate(
    gender_adj = factor(gender,
                        levels = c(0, 1, 2),
                        labels = c("Men", "Women", "Others")),
    sexorient_adj = factor(sexorient,
                           levels = c(0, 1, 2),
                           labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed")),
    age = ifelse(nchar(as.character(age)) > 3, NA, age)
  )

stopvape_relapsed

```


# prepare data
```{r}
# Convert categorical variables to a factor with proper labels
stopvape_relapsed$gender_adj <- factor(stopvape_relapsed$gender,
                            levels = c(0, 1, 2),
                            labels = c("Men", "Women", "Others"))
stopvape_relapsed$sexorient_adj <- factor(stopvape_relapsed$sexorient,
                               levels = c(0, 1, 2),
                               labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed"))
# 
# stopvape_relapsed <- stopvape_relapsed %>%
#   mutate(age = ifelse(nchar(as.character(age)) > 3, NA, age))


summary(stopvape_relapsed$age)
```

```{r}
#split into groups low and high quitting confidence 
stopvape_relapsed$conf_group <- cut(stopvape_relapsed$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right = TRUE)


# Filter to unique user IDs, create confidence groups, and get descriptive statistics
stopvape_relapsed %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  tbl_summary(
    by = conf_group,  # Group by confidence level
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%        # Add p-values for group comparisons
  add_overall()      # Add overall column
```


```{r}
(v <- names(stopvape_relapsed)[c(2, 14, 32, 33, 43, 44, 42, 12, 13, 28, 40, 41)])
set.seed(3098) 

# stopvape_relapsed.mice  <- mice(stopvape_relapsed,
#                   seed  = 3098,
#                   m     = 5,
#                   print = F)
stopvape_relapsed.mice <- mice(stopvape_relapsed[,v], method = "cart") # large number of unbalanced factor variables, use cart method
stopvape_relapsed.mice$method
plot(stopvape_relapsed.mice)

densityplot(stopvape_relapsed.mice) # for continuous variables
densityplot(stopvape_relapsed.mice, ~ ecusv_3|.imp) #look at self-confidence by imputation

```

```{r}
fit1_relapsed <- with(
  stopvape_relapsed.mice,
  coxph(Surv(Time.elapsed, status) ~ ecusv_3 + cluster(userid))
)

fit2_relapsed <- with(
  stopvape_relapsed.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 + age + country + gender_adj +
          sexorient_adj + race + ecu17_1 +
          ecusv_2 + mhealth1_1 +
          cluster(userid))
)

fit3_relapsed <- with(
  stopvape_relapsed.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 * gender_adj +
          cluster(userid))
)

fitdemo_relapsed <- with(
  stopvape_relapsed.mice,
  coxph(Surv(Time.elapsed, status) ~
          ecusv_3 + age + country +
          gender_adj + sexorient_adj + race +
          cluster(userid))
)

pooled_fit1_relapsed <- pool(fit1_relapsed)
pooled_fit2_relapsed <- pool(fit2_relapsed)
pooled_fit3_relapsed <- pool(fit3_relapsed)
pooled_fitdemo_relapsed <- pool(fitdemo_relapsed)

summary(pooled_fit1_relapsed, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fit2_relapsed, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fit3_relapsed, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_fitdemo_relapsed, conf.int = TRUE, exponentiate = TRUE)
```

```{r}

# return all imputed datasets combined into one big data frame
stopvape_relapsed.pooled <- complete(stopvape_relapsed.mice, method = "long") 
summary (stopvape_relapsed.pooled)
nrow(stopvape_relapsed.pooled)

surv_elapsed <- Surv(time = stopvape_relapsed.pooled$Time.elapsed, 
                 event = stopvape_relapsed.pooled$status)


fit1_relapsed <- coxph(surv_elapsed~ecusv_3,cluster(stopvape_relapsed.pooled$userid),data=stopvape_relapsed.pooled)
fit2_relapsed <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race+ecu17_1+ecusv_2+mhealth1_1,cluster(stopvape_relapsed.pooled$userid),data=stopvape_relapsed.pooled)
fit3_relapsed <- coxph(surv_elapsed~ecusv_3*gender_adj,cluster(stopvape_relapsed.pooled$userid),data=stopvape_relapsed.pooled)

fitdemo_relapsed <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race,cluster(stopvape_relapsed.pooled$userid),data=stopvape_relapsed.pooled)

fit1_relapsed
fit2_relapsed
fit3_relapsed
fitdemo_relapsed

#95% CI
exp(confint(fit1_relapsed))
exp(confint(fit2_relapsed))
exp(confint(fit3_relapsed))
exp(confint(fitdemo_relapsed))

#Run AIC to compare predictive performance of models
AIC(fit1_relapsed)
AIC(fit2_relapsed)
AIC(fit3_relapsed)
AIC(fitdemo_relapsed)

```

