---
title: "StopVaping_SelfConfidence_submission"
author: "Jessie Tian"
date: "2025-12-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(rms)
library(lattice)
library(dplyr)
library(tidyr)
library(survival)
library(ggplot2)
library(skimr)
library(naniar)
library(survival)
library(survminer)
library(lubridate)
library(mice)
library(gtsummary)
library(dplyr)

require("knitr")
```

```{r}

setwd("~/Desktop/OTRU/Quit Vaping App ")
stopvape <- read_csv("unimputed_dataset_allages.csv", col_names = TRUE)

stopvape.dd <- datadist(stopvape)
options(datadist="stopvape.dd")

stopvape_complete <- na.omit(stopvape)
sum(is.na(stopvape_complete))


#Complete cases 
nrow(stopvape[!complete.cases(stopvape), ])/nrow(stopvape)*100
#Incomplete cases 17.82%
stopvape_comp <- stopvape[complete.cases(stopvape),]
# df_comp$wave <- NULL
#n=5,288, p=37



```


ecusv_3: Confidence in quitting vaping within next month on a scale of 1-10
ecu17_1: Intention to quit at baseline (0= yes, but within next 6 months or beyond 6 months, 1= within next month,)
ecusv_2: Past-year quit attempts (0= no, 1=yes)
mhealth1_1: Perceived mental heath status (0= fair or poor, 1= good, very good, or excellent)

# prepare data
```{r}
# Convert categorical variables to a factor with proper labels
stopvape$gender_adj <- factor(stopvape$gender,
                            levels = c(0, 1, 2),
                            labels = c("Men", "Women", "Others"))
stopvape$sexorient_adj <- factor(stopvape$sexorient,
                               levels = c(0, 1, 2),
                               labels = c("Heterosexual", "LGBTQ", "Asexual/Undisclosed"))

stopvape <- stopvape %>%
  mutate(age = ifelse(nchar(as.character(age)) > 3, NA, age))


summary(stopvape$age)
```

# Descriptive table
```{r}

#split into groups low and high quitting confidence 
stopvape$conf_group <- cut(stopvape$ecusv_3,
                           breaks = c(0, 5, 10),
                           labels = c("Low", "High"),
                           right = TRUE)

# Get descriptive statistics for unique user IDs only
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n}  ({p}%)"
    )
  )

# Filter to unique user IDs, create confidence groups, and get descriptive statistics
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%  # Keep only unique userids
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  tbl_summary(
    by = conf_group,  # Group by confidence level
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%        # Add p-values for group comparisons
  add_overall()      # Add overall column


# Select specific variables and group by confidence
stopvape %>%
  distinct(userid, .keep_all = TRUE) %>%
  mutate(conf_group = cut(ecusv_3,
                         breaks = c(0, 5, 10),
                         labels = c("Low", "High"),
                         right = TRUE)) %>%
  select(age,gender,sexorient,race,status, Time.elapsed, ecu17_1, ecusv_2, ecusv_3, mhealth1_1, conf_group) %>%
  tbl_summary(
    by = conf_group,
    statistic = list(
      all_continuous() ~ "{median} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%
  add_overall()

# Calculate statistics using unique user_id
quit_stats <- stopvape %>%
  group_by(userid) %>%
  summarise(
    quit_attempts = n(),  # Count records per user (each record = 1 attempt)
    .groups = 'drop'
  )

# Calculate final statistics
final_stats <- quit_stats %>%
  summarise(
    # Average number of quit attempts per participant
    avg_quit_attempts = mean(quit_attempts),
    # Number of participants with only one challenge
    participants_one_challenge = sum(quit_attempts == 1),
    # Maximum number of challenges completed
    max_challenges = max(quit_attempts)
  )

print(final_stats)


# Count how many times each userid appears (number of challenges per user)
user_challenge_counts <- table(stopvape$userid)

```

# Missing data and multiple imputations (pattern of missingness)
```{r}

skim(stopvape)
summary(stopvape, na.rm = TRUE) 

# Explore patterns of missing data
na.pat <- naclus(stopvape[,-1])
naplot(na.pat)

vis_miss(stopvape) #display entire dataset to find missing, if sort by time, can see if time dependent missingness throughout study
gg_miss_var(stopvape) # give number missing sorted
gg_miss_var(stopvape, show_pct = TRUE) # shown as percentage
gg_miss_case(stopvape)

gg_miss_upset(stopvape) # see what is missing when other variables are missing
n_var_miss(stopvape) #max number missing is 23 variables

ggplot(
  data = stopvape,
  mapping = aes(x = Time.elapsed, y = ecusv_3)) +     
  geom_miss_point()

stopvape$age[stopvape$age== 5198023886] <- NA

### Generate 5 imputed data sets

(v <- names(stopvape)[c(14, 32, 33, 43, 44, 42, 12, 13, 28, 40, 41)])

# md.pattern(stopvape[,v], rotate.names = TRUE)

set.seed(3098) 
stopvape.mice <- mice(stopvape[,v], method = "cart") # large number of unbalanced factor variables, use cart method 
stopvape.mice$method
plot(stopvape.mice)

densityplot(stopvape.mice) # for continuous variables
densityplot(stopvape.mice, ~ ecusv_3|.imp) #look at self-confidence by imputation


```

# Run Cox proportional hazard model for unadjusted, adjusted, and gender*self-confidence interaction on pooled imputated data
```{r}

# return all imputed datasets combined into one big data frame
stopvape.pooled <- complete(stopvape.mice, method = "long") 

fit1 <- coxph(surv_elapsed~ecusv_3,cluster(stopvape.pooled$userid),data=stopvape.pooled)
fit2 <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race+ecu17_1+ecusv_2+mhealth1_1,cluster(stopvape.pooled$userid),data=stopvape.pooled)
fit3 <- coxph(surv_elapsed~ecusv_3*gender_adj,cluster(stopvape.pooled$userid),data=stopvape.pooled)

fitdemo <- coxph(surv_elapsed~ecusv_3+age+country+gender_adj+sexorient_adj+race,cluster(stopvape.pooled$userid),data=stopvape.pooled)

#95% CI
exp(confint(fit1))
exp(confint(fit2))
exp(confint(fit3))
exp(confint(fitdemo))

#Run AIC to compare predictive performance of models
AIC(fit1)
AIC(fit2)
AIC(fit3)
AIC(fitdemo)

#Stratified cox model by gender across imputations
# men unadjusted 
cox_mi_men_unadj <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Men"))
)
pooled_men_unadj <- pool(cox_mi_men_unadj)

# Women adjusted
cox_mi_women_unadj <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Women"))
)
pooled_women_unadj <- pool(cox_mi_women_unadj)

# others unadjusted 
cox_mi_others_unadj <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Others"))
)
pooled_others_unadj <- pool(cox_mi_others_unadj)
  
# Men adjusted
cox_mi_men <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3 + age + country + sexorient_adj + race + ecu17_1 + ecusv_2 + mhealth1_1,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Men"))
)
pooled_men <- pool(cox_mi_men)
# Women adjusted
cox_mi_women <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3 + age + country + sexorient_adj + race + ecu17_1 + ecusv_2 + mhealth1_1,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Women"))
)
pooled_women <- pool(cox_mi_women)
# Others adjusted 
cox_mi_others <- with(
  data = stopvape.mice,
  expr = coxph(Surv(Time.elapsed, status) ~ ecusv_3 + age + country + sexorient_adj + race + ecu17_1 + ecusv_2 + mhealth1_1,
               cluster = stopvape.pooled$userid,
               subset = (gender_adj == "Others"))
)
pooled_others <- pool(cox_mi_others)


summary(pooled_men_unadj, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_women_unadj, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_others_unadj, conf.int = TRUE, exponentiate = TRUE)

summary(pooled_men, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_women, conf.int = TRUE, exponentiate = TRUE)
summary(pooled_others, conf.int = TRUE, exponentiate = TRUE)
```

# Univariable Kaplan-Meier curve of vaping relapse survival by quitting confidence levels (low, high)

```{r}

surv_elapsed <- Surv(time = stopvape.pooled$Time.elapsed, 
                 event = stopvape.pooled$status)
relapse.km <- npsurv(surv_elapsed ~ conf_group, data = stopvape.pooled)

survplotp(relapse.km,
          conf.int = TRUE,
          col = c("red", "blue"),
          lty = 1:2,
          xlim = c(0,45),
          time.inc = 2,
          xlab = "Time Elapsed (days)",
          ylab = "Proportion Remaining Vape-Free",
          main = "Survival by Confidence in Quitting (Low vs High)",
          label.curves = list(
            title = "Quitting self-efficacy level"
          )
)



#median survival time for each quitting confidence level 1-10
survfit(Surv(Time.elapsed,status)~ecusv_3, data = stopvape.pooled)
survfit(surv_elapsed ~ conf_group, data = stopvape.pooled)


# Separate data by sex (adjust variable name if needed)
stopvape_male <- stopvape.pooled[stopvape.pooled$gender_adj == "Men", ] 
stopvape_female <-stopvape.pooled[stopvape.pooled$gender_adj == "Women", ]
stopvape_others <- stopvape.pooled[stopvape.pooled$gender_adj == "Others", ]

surv_male <- Surv(time = stopvape_male$Time.elapsed, event = stopvape_male$status)
surv_female <- Surv(time = stopvape_female$Time.elapsed, event = stopvape_female$status)
surv_others <- Surv(time = stopvape_others$Time.elapsed, event = stopvape_others$status)

relapse.km.male <- npsurv(surv_male ~ conf_group, data = stopvape_male)
relapse.km.female <- npsurv(surv_female ~ conf_group, data = stopvape_female)
relapse.km.others <- npsurv(surv_others~ conf_group, data = stopvape_others)

# Create side-by-side plots
par(mfrow = c(1, 2))

survplotp(relapse.km.male, 
          conf.int = TRUE, 
          col = c("red", "blue"), 
          lty = 1:2,
          xlim = c(0,30),
          time.inc = 2,
          xlab = "Time Elapsed for Men",
          ylab = "Proportion Remaining Vape-Free",
          main = "Survival by Confidence - Men")

survplotp(relapse.km.female, 
          conf.int = TRUE, 
          col = c("red", "blue"), 
          lty = 1:2,
          xlim = c(0,30),
          time.inc = 2,
          xlab = "Time Elapsed for Women",
          ylab = "Proportion Remaining Vape-Free",
          main = "Survival by Confidence - Women")

survplotp(relapse.km.others, 
          conf.int = TRUE, 
          col = c("red", "blue"), 
          lty = 1:2,
          xlim = c(0,30),
          time.inc = 2,
          xlab = "Time Elapsed for Others",
          ylab = "Proportion Remaining Vape-Free",
          main = "Survival by Confidence - Others")
```
```{r}

png("survival_plot.png", width = 1800, height = 1400, res = 400)

survplot(
  relapse.km,
  conf.int = TRUE,
  col = c("red", "blue"),
  lty = c(1, 1),
  lwd = 2,
  xlim = c(0, 45),
  time.inc = 2,
  xlab = "Time Elapsed (days)",
  ylab = "Proportion Remaining Vape-Free",
  main = "Survival by Confidence in Quitting (Low vs High)",
  label.curves = FALSE
)

dev.off()



```

### log-rank MH test → highly significant difference between the survival curves of the groups
```{r}
#log-rank MH test 
#compute the test of the difference between the survival curves
sur.mh <- survdiff(surv_elapsed ~ conf_group, data = stopvape)
sur.mh 

```



